Wikipedia Text Mining and Entropy Analysis

### 2019-01-24 cogsci/position-info.Rmd
rewrites

---
title: "Rate of Information "
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

abstract: 
    "A major line of research in the cognitive sciences has asked how speakers and writers distribute information in the linguistic material they produce [include citations here or rewrite this paragraph]. We provide evidence against the Uniform Information Density hypothesis (Jaeger, 2010) which proposes that information is transmitted at a constant rate close to channel capacity in human communication [Why does it matter that this is the hypothesis we are arguing against]. We instead argue that the information structure and transmission rate in human communication varies predictably from language to language based on phonological and syntactic features. We draw evidence from written and spoken corpora, including corpora of parent and child communication at various ages and large-scale online corpora from Wikipedia in a typologically diverse collection of languages [Rewrite final sentence for clarity]."

@@

Yu et al. (2016) challenges the UID hypothesis through examining the distribution of uncertainty across sentence positions within the text portion of the English-language British National Corpus (BNC). They partition the corpus by sentence length in number of words, and proceed for each sentence length $k$ within the corpus. for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e. the number of sentences of length $k$.

$$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
With this measure, Yu et al. compute the unigram entropy at each position of sentences of each length within the corpus. The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances

Let us assume that the UID perspective is correct. Then entropy should monotonically and uniformly increase at all word positions across utterances on average. We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.

# Methods

For spoken data, we choose to examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages. We expect based on the UID account that the entropy curves of the children will be significantly lower than those of the adults, due to the children's comparatively lower channel capacity in information transmission [more support]. We first used the Brown and Providence English corpora from CHILDES. The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes. The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home. We divided each corpus by speaker into child and non-child categories. We further divided the corpora by utterance length, so that all sentences of length $k$ (e.g. $6$) were grouped together. Finally, within each utterance length, we computed the unigram entropy measure for each position.

@@

We also ran this analysis on Spanish, German, Mandarin and Japanese corpora from CHILDES. We used the Shiro corpus for Spanish (Shiro, 2000), which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds. The German corpus we used was the Wagner corpus of spontaneous parent-child speech in the home, collected in single sessions for children between 1 and 14 years of age (Wagner, 1985). We used the Zhou dinner corpus for Mandarin Chinese (Li & Zhou, 2015), which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai. We used the Okayama corpus for Japanese (Okayama, 1973), which consists of conversations between 2 to 4-year-old children and their mothers in single sessions.

For each corpus, we accessed transcripts of the corpus provided through the TalkBank system and computed over Roman alphabet transcriptions or transliterations of the original transcriptions. For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries, and for Japanese we used romanji (Roman alphabet) transliterations of words in the corpus. Japanese Hiragana, Katakana and Kanji writing systems, and the Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Japanese and Chinese writing including spaces between word boundaries can have a negative effect on reading times (Sainio et al, 2007; Bai et al, 2008).

@@

```{r wagner_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Wagner corpus unigram entropy"}
img <- png::readPNG("figs/Spanish_unigram_entropies_with_stemming.png")
grid::grid.raster(img)
```

@@

```{r okayama_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Okayama corpus unigram entropy"}
img <- png::readPNG("figs/Japanese_okayama_entropy.png")
grid::grid.raster(img)
```

# Results & Analysis
The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce . This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees. We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve. This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.

We found a distinct three-step distribution for English, Spanish and German CHILDES corpora, with a slight dip in the penultimate position of each sentence. We attribute the penultimate dip in these unigram entropy curves to the fact that most of the utterances in the CHILDES English, Spanish and German corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances. The beginnings of utterances in the English, German and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position. The final position of utterances in child-directed speech is known to be important, dating back to Aslin (1993). The Mandarin and Japanese data, by comparison, have noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions. The Japanese entropy curve begins high and then drops, unlike all of the other entropy curves.

```{r MI_wagner, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Mutual Information measures on utterances of length 6 in the Wagner corpus"}
img <- png::readPNG("figs/wagner_mi_6.png")
grid::grid.raster(img)
```

```{r MI_okayama, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Mutual Information measures on utterances of length 6 in the Okayama corpus"}
img <- png::readPNG("figs/okayama_mi_6.png")
grid::grid.raster(img)
```

The mutual information statistic is a symmetric measure of the information gained about one random variable by learning about another, a bigram measure and indicator of the effect than the context of a word can have on its unigram entropy value. As seen in the images below, the final word in utterances of length $6$ in German has a significant mutual information value with all other words in the utterance on average. This means that the words appearing in positions $1$ through $5$ all play a role in determining the word that appears in position $6$ and vice versa. By contrast, the words that appear in positions $1$ through $4$ have little influence in determining the word that appears in position $5$, the penultimate word that is often a determiner or function word, and instead the final word has the greatest influence in determining the penultimate word. For Japanese, the final word in an utterance has little to no influence on determining the rest of the words and vice versa, whereas the other words all play in role in determining what will go in each other's positions.

# Large-scale text data [Why does this comparison matter?]
We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES. Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language. Each language corpus was cleaned and limited to sentences between $6$ and $50$ words. Similar to the data from CHILDES, we divided each corpus by sentence length, and then computed the unigram entropy measure on each word position within each sentence length.

@@

## Clustering
In the *absolute* and *relative 5* treatments, each language is cast as a point in $5$-dimensional space. In the *relative 10* treatment, each language is cast as a point in $10$-dimensional space. This allows for direct comparison using cosine similarity and moreover for unsupervised clustering analysis to compare the results of the Wikipedia entropy analysis to known typological features within the langauges in the Wikipedia dataset. Using the R hclust package, we performed a hierarchical clustering of the Wikipedia slope data within each treatment. A subset of the results at a glance for the *absolute* treatment are below.

```{r absolute_clusters, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Absolute clustering of Wikipedia corpora"}
img <- png::readPNG("figs/absolute_clusters.png")
grid::grid.raster(img)
```

## Language Features

To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we looked at the list of $144$ linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013). We limited the languages in the WALS database to only those in our Wikipedia dataset and performed missing-value imputation to obtain the features not coded in WALS for the languages from Wikipedia. We then performed k-modes clustering and compared the results with the hierarchical clustering from Wikipedia. [Still working on this]

We also checked the effects of individual features on the embeddings of languages in the different treatments. [Still working on this]

>>>>

---
title: "Information is Not Communicated Uniformly: Evidence from Spoken and Written Corpora"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

abstract: 
    "We provide evidence against the popular Uniform Information Density hypothesis (Levy & Jaeger, 2007) which proposes that information is transmitted at a constant rate close to channel capacity in human communication. Using a method based on the original Genzel & Charniak (2002) entropy measure, we construct a word-level model for entropy applicable to both spoken and written corpora. We apply this model to corpora from Wikipedia in well over a hundred languages. We find that not only is the Uniform Information Density hypothesis prediction wrong, but also that the by-word entropy distribution of a language is related to the typological features of the language. We use this evidence to suggest that people do not communicate information at a uniform rate, but that information distribution varies from language to language based on phonogical, morphological and syntactic features."

@@

What Zhan and Levy (2018) and most work in favor of UID have in common is they focus on information distribution at particular points in a sentence. By contrast, Jain et al. (2018) examines word order across sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order. For this paper, we focus on the sentence level instead of specific points, following a method described in Yu et al. (2016).

Yu et al. (2016) challenges the UID hypothesis through examining the distribution of uncertainty across sentence positions within the text portion of the English-language British National Corpus (BNC). They partition the corpus by sentence length in number of words, and proceed for each sentence length $k$ within the corpus. for each word position $X$ of sentences of length $X$, let each word $w$ be a unique word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position $X$ divided by the number of total words that occur in position $X$ i.e. the number of sentences of length $k$. Then $p(w)$ is the probability of randomly obtaining word $w$ by choosing a word in position $X$ in sentences of length $k$.

$$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
With this measure, Yu et al. compute the unigram entropy at each position of sentences of each length within the corpus. The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances. Genzel & Charniak (2002) similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus. UID applies this uniformity of entropy rate in sentences to all levels of speech, and so the Yu et al. method, which examines text at the word level, should find an affine function at the word level.

Let us assume that the UID perspective is correct. Then we expect that the entropy curve for every language should be close to an affine, linear function with some noise. Entropy should monotonically and uniformly increase at all word positions across utterances on average. We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.

# Spoken data

## Methods

We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of telephone conversations [CITATION].

We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages. Children are not fully developed speakers, and so we expect to observe a different shape of the distribution. We first used the Brown and Providence English corpora from CHILDES. The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes. The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home. We divided each corpus by speaker into child and non-child categories. We further divided the corpora by utterance length, so that all sentences of length $k$ (e.g. $6$) were grouped together. Finally, within each utterance length, we computed the unigram entropy measure for each position.

@@

We also ran this analysis on Spanish and Mandarin corpora from CHILDES. We used the Shiro corpus for Spanish (Shiro, 2000), which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds. We used the Zhou dinner corpus for Mandarin Chinese (Li & Zhou, 2015), which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.

For each corpus, we accessed transcripts of the corpus provided through the TalkBank system and computed over Roman alphabet transcriptions or transliterations of the original transcriptions. For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries, and for Japanese we used romanji (Roman alphabet) transliterations of words in the corpus. The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times (Bai et al, 2008).

@@

@@

## Results & Analysis
The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce . This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees. We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve. This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.

We found a distinct three-step distribution for English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence. We attribute the penultimate dip in these unigram entropy curves to the fact that most of the utterances in the CHILDES English and Spanish corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances. The beginnings of utterances in the English and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position. The final position of utterances in child-directed speech is known to be important, dating back to Aslin (1993). The Mandarin corpus entropy curve, by comparison, has a noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.

# Written data
We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES. WHY ARE WE DOING THIS.

## Methods

Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language. Each language corpus was cleaned and limited to sentences between $6$ and $50$ words. Similar to the data from CHILDES, we divided each corpus by sentence length, and then computed the unigram entropy measure on each word position within each sentence length.

@@

## Results and Analysis

In the *absolute* and *relative 5* treatments, each language is cast as a point in $5$-dimensional space. In the *relative 10* treatment, each language is cast as a point in $10$-dimensional space. This allows for direct comparison using cosine similarity and moreover for unsupervised clustering analysis to compare the results of the Wikipedia entropy analysis to known typological features within the langauges in the Wikipedia dataset. Using the R hclust package, we performed a hierarchical clustering of the Wikipedia slope data within each treatment. A subset of the results at a glance for the *absolute* treatment are below.

To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we looked at the list of $144$ linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013). We limited the languages in the WALS database to only those in our Wikipedia dataset and performed missing-value imputation to obtain the features not coded in WALS for the languages from Wikipedia. [Miss MDA or do another kind of missing value imputation, description of method]

One approach is to cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971). However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis. This is a computationally intractable operation. Additionallly, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning.

We instead check the effects of individual features on the embeddings of languages in the different treatments. We computed pairwise cosine similarity between each pair of language vectors within a treatment. For a subset of WALS features which had values entered for most of the languages we obtained from Wikipedia, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature. We found this to be true.

RESULTS OF GLM BELOW.

### 2019-01-24 wiki_entropy/wiki_clustering.Rmd
rewrites

```{r}
cosines <- wkpd %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, cosine, -language1)

@@

wf_comp <- wf_sub %$%
  data %>%
  do.call(cbind, args=.)

wf_comp <- wf_comp[, -seq(3, length(wf_comp), 2)] 
langs_to_compare <- wf_comp$language
wf_comp %<>% 
  as.data.frame(row.names=wf_comp$language)

wkpd_comp <- wkpd %>%
    filter(Language %in% langs_to_compare) 
wkpd_comp %<>% as.data.frame(row.names=wkpd_comp$Language)

lang_pairs <- langs_to_compare %>%
  combn(2, simplify = FALSE)

K <- length(lang_pairs)

lang_x <- lang_pairs %>%
    unlist() %>%
    matrix(nrow = K, byrow=TRUE) %>%
    as.data.frame() %>%
    mutate(wiki = lsa::cosine(wkpd_comp[V1, ], wkpd_comp[V2, ]), 
           wals = lsa::cosine(wf_comp[V1, ], wf_comp[V2, ]))

>>>>

```{r}
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

cosines <- wkpd %>%
  column_to_rownames("Language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, cosine, -language1)

@@
